---
title: "Mesophotic.org database release 2.1"
author: "Gonzalo Pérez-Rosales, https://github.com/gonzaloprb"
date: "14/08/2024"
output: html_document
---

### Load the following packages for manipulating and viewing the data:
```{r message=FALSE}
require(dplyr)
require(tidyr)
require(ggplot2)
require(ggpubr)
require(tidyverse)
require(cowplot)
library(RcppRoll)
require (plyr); 
require(reshape2)
require (patchwork)
rm(list=ls()) 
```


## Import the database directly from the website into R

Check its dimensions (rows x columns), and list the available fields/columns:
```{r}

# first database release used
publs <- read.csv("http://www.mesophotic.org/publications.csv", as.is=TRUE)
# Update to the Github Directory
setwd("~/Documents/AAASea_Science/Mesophotic.org/")

publs[sapply(publs, is.character)] <- lapply(publs[sapply(publs, is.character)], as.factor)

# when need to use year as factor
publs$year <- factor(publs$publication_year)

# Quick check up and info of how the data is structured
dim(publs)
colnames(publs)
levels(publs$publication_format)

# Summary format
publs %>% 
  group_by(publication_format) %>% 
   dplyr::summarise(count=n())

levels(publs$publication_type)
# Publication type
publs %>% 
  group_by(publication_type) %>% 
   dplyr::summarise(count=n())

# Complete publications columns
publs_complete <- publs %>% complete(publication_type,publication_year, focusgroups,fill = list(publication_year = NA))

```


## This code makes the figures used in the manuscript. 
(At the bottom of the R pipeline there are many other figure outputs, which are not used in the Manuscript)

Plot by platforms with max Depth
Make a box plot.
- filtering publications with a original data, validated, stats and a single platform only, 
- Consider max depths
- Please note that for SCUBA (open-circuit or unspecified), there are 10 publications deeper than 100 m and 25 deeper than 70 m

```{r}
publs <- subset(publs, publication_type == "scientific" & publication_format == "article")
publs_data <- publs[c("id","validated","included_in_stats","publication_year","original_data", "platforms","min_depth", "max_depth")]
publs_data <- subset (publs_data, original_data == "true")
publs_data <- subset (publs_data, validated == "true")
publs_data <- subset (publs_data, included_in_stats == "true")

# Filter to publications having a single platform
# Made that by deleting rows containing ";"
platforms_df <- publs_data[!grepl(";", publs_data$platforms),]
# Delete columns where platforms is empty
platforms_df <- platforms_df[-which(platforms_df$platforms == ""), ]

# Remove publications where max_depth == 0 (No depth range or only mesophotic depth range)
platforms_df <- platforms_df %>% filter(max_depth > 29) # Dropped publications where max_depth less than 29
# To consider the mesophotic depth range 


# Transform publications where min_depth is NA to Max_Depth
# How many do we have?
sum(is.na(platforms_df$min_depth))
# Replace by max depth
platforms_df$min_depth <- ifelse(is.na(platforms_df$min_depth), platforms_df$max_depth, platforms_df$min_depth)
# check
sum(is.na(platforms_df$max_depth)) 

# Filter publications deeper than Mesophotic for the min_depth 
platforms_df <- platforms_df %>% filter(min_depth < 151) # Keep publications where min_depth is less than 150 m. Otherwise, outside Mesophotic range

# Make the plot - with all platforms (see filtered plot after)
ggplot(data = platforms_df, aes (y=max_depth, x = factor(platforms))) +   
  geom_boxplot() +
  labs(y = "Depth (m)", x = "", title = "") + scale_y_reverse() +
  theme_bw() + scale_x_discrete(position = "top") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Keep only certain platforms and arrange by depth
keep_plat <-  c ("SCUBA (open-circuit or unspecified)","Autonomous Underwater Vehicle (AUV)","Rebreather","Baited Remote Underwater Video (BRUV)","Fishing","Remotely Operated Vehicle (ROV)","Manned Submersible", "Dredging / trawling", "Sonar / Multibeam")
platforms_df <- subset(platforms_df,platforms %in% keep_plat)

# Transform characters to simplified names
platforms_df$platforms<- gsub("SCUBA \\(open-circuit or unspecified\\)", "SCUBA",platforms_df$platforms)
platforms_df$platforms<- gsub("Autonomous Underwater Vehicle \\(AUV\\)", "AUV",platforms_df$platforms)
platforms_df$platforms<- gsub("Sonar / Multibeam", "Sonar",platforms_df$platforms)
platforms_df$platforms<- gsub("Baited Remote Underwater Video \\(BRUV\\)", "BRUV",platforms_df$platforms)
platforms_df$platforms<- gsub("Dredging / trawling", "Dredging",platforms_df$platforms)
platforms_df$platforms<- gsub("Remotely Operated Vehicle \\(ROV\\)", "ROV",platforms_df$platforms)
platforms_df$platforms<- gsub("Manned Submersible", "Submersible",platforms_df$platforms)

# Set in the right order
platforms_df$platforms <- factor(platforms_df$platforms, levels = c("SCUBA","Submersible","Rebreather","ROV","Dredging","Fishing","BRUV","AUV","Sonar"))


# Make a summary table with the number of publications per category: 
pubs_platforms <- ddply(platforms_df, ~ platforms,
                   function(x){c(Pubs_number= length(x$id))})

pubs_platforms_depth <- ddply(platforms_df, ~ platforms,
                   function(x){c(Max_depth_mean= mean(x$max_depth))})

pubs_platforms_depth_median <- ddply(platforms_df, ~ platforms,
                   function(x){c(Max_depth_median= median(x$max_depth))})
# Merge dataframes
pubs_platforms <- merge (pubs_platforms,pubs_platforms_depth)
pubs_platforms <- merge (pubs_platforms,pubs_platforms_depth_median)


# Function to give number of publications 
give.n <- function(x){
  return(c(y = median(x)*1.05, label = length(x))) 
  # experiment with the multiplier to find the perfect position
}

# Make a first boxplot. Single platform, not duplicate if multilocation, remove jitter, hard stop of plot at 200 m
ggplot(data = platforms_df, aes (y=max_depth, x = factor(platforms))) + # geom_point() +
  geom_boxplot(outlier.shape = NA) + stat_summary(fun.data = give.n, geom = "text", size=5, 
                                mapping = aes(y = 20), vjust = -0.5, fun.y = median,
                                position = position_dodge(width = 0.75)) + 
  labs(y ="Max depth (m)", x = "", title = "") + 
  scale_y_reverse() + coord_cartesian(ylim=c(300, 0)) +
  scale_y_reverse(breaks = c(300, 250, 200, 150, 100, 50, 0)) + # Otherwise, scale_y_reverse(limits=c(4000,0))
  theme_bw() + scale_x_discrete(position = "top") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, lineheight=1))


##### Make the Figure 3 (a) / Fig_Platform_MaxDepth
# Violin plot 
Fig_Platform_MaxDepth <- ggplot(data = platforms_df, aes (y=max_depth, x = factor(platforms))) + # geom_point() +
  geom_violin(draw_quantiles = c(0.5),trim=T, width = 2) +
    #geom_violin(draw_quantiles = c(0.1, 0.5, 0.9),trim=F, width = 2) +
  geom_jitter(size = 0.5, position = position_jitter(seed = 1, width = 0.005)) +
  stat_summary(fun.data = give.n, 
    geom = "text", size=5, mapping = aes(y = 15), vjust = -0.5, fun.y = median,
    position = position_dodge(width = 0.75))+
  labs(y ="Max depth (m)", x = "", title = "Direct human presence                      Indirect / collection                Observations only") +
  scale_y_reverse() + coord_cartesian(ylim=c(300, 0)) +
  scale_y_reverse(breaks = c(300, 250, 200, 150, 100, 50, 0)) + # Otherwise, scale_y_reverse(limits=c(4000,0))
  theme_bw() + scale_x_discrete(position = "top") + 
  theme(axis.text.x = element_text(size = 10))
Fig_Platform_MaxDepth

# Please note that for SCUBA (open-circuit or unspecified), there are 10 publications deeper than 100 m and 25 deeper than 70 m

ggsave ( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Fig_Platform_MaxDepth.pdf", Fig_Platform_MaxDepth,width = 8.5, height = 3.5)
ggsave ( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Fig_Platform_MaxDepth.jpeg", Fig_Platform_MaxDepth,width = 8.5, height = 3.5)
ggsave ( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Fig_Platform_MaxDepth.svg", Fig_Platform_MaxDepth,width = 8.5, height = 3.5)


rm (top_locations,pubs_platforms_pubs_ecosystem, pubs_platforms_depth_median, pubs_platforms_locationspubls, order, Locations, publs_complete, pubs_platforms_depth_median, platforms_df, pubs_platforms_depth, Fig_Platform_MaxDepth)

```


Figure of research fields by platforms 
These plots also exist by focus groups below

- filtering publications with a original data, validated, stats and a single platform only, 

```{r}
publs <- subset(publs, publication_type == "scientific" & publication_format == "article")
publs_data <- publs[c("id","validated","included_in_stats","publication_year","original_data", "platforms","min_depth", "max_depth", "fields")]
publs_data <- subset (publs_data, original_data == "true")
publs_data <- subset (publs_data, validated == "true")
publs_data <- subset (publs_data, included_in_stats == "true")

# Generate a new database
platforms_df <- publs_data

# Work on PLATFORMS
# Here we Keep all platforms! We do not filter to publications with single platform! 
# First delete columns where platforms is empty
platforms_df <- platforms_df[-which(platforms_df$platforms == ""), ]

# Work on FIELDS
# Remove rows with no associated `fields`
platforms_df <- platforms_df[!(is.na(platforms_df$fields) | platforms_df$fields == ""), ]

# Remove publications where max_depth == 0
platforms_df <- platforms_df %>% filter(max_depth > 29) # Dropped publications where max_depth > 29

# Transform publications where min_depth is NA to Max_Depth
# How many do we have?
sum(is.na(platforms_df$min_depth))
# Replace by max depth
platforms_df$min_depth <- ifelse(is.na(platforms_df$min_depth), platforms_df$max_depth, platforms_df$min_depth)
# check
sum(is.na(platforms_df$max_depth)) 

# Filter publications deeper than Mesophotic for the min_depth
platforms_df <- platforms_df %>% filter(min_depth < 151) # Keep publications where min_depth is less than 150 m
# To consider the Mesophotic depth range

# Necessary to separate by rows. Be aware that these commands generate more publications numbers due to multiplatform and multifields - Keep all of them but as separate rows 
platforms_df <- separate_rows(platforms_df, "platforms", sep=";\\s+") 
platforms_df <- separate_rows(platforms_df, "fields", sep=";\\s+") 
# We keep all rows. No need to delete duplicated

# Keep only certain platforms and arrange by depth
keep_plat <-  c ("SCUBA (open-circuit or unspecified)","Autonomous Underwater Vehicle (AUV)","Rebreather","Baited Remote Underwater Video (BRUV)","Fishing","Remotely Operated Vehicle (ROV)","Manned Submersible", "Dredging / trawling", "Sonar / Multibeam")
platforms_df <- subset(platforms_df,platforms %in% keep_plat)

# Transform characters to simplified names
platforms_df$platforms<- gsub("SCUBA \\(open-circuit or unspecified\\)", "SCUBA",platforms_df$platforms)
platforms_df$platforms<- gsub("Autonomous Underwater Vehicle \\(AUV\\)", "AUV",platforms_df$platforms)
platforms_df$platforms<- gsub("Sonar / Multibeam", "Sonar",platforms_df$platforms)
platforms_df$platforms<- gsub("Baited Remote Underwater Video \\(BRUV\\)", "BRUV",platforms_df$platforms)
platforms_df$platforms<- gsub("Dredging / trawling", "Dredging",platforms_df$platforms)
platforms_df$platforms<- gsub("Remotely Operated Vehicle \\(ROV\\)", "ROV",platforms_df$platforms)
platforms_df$platforms<- gsub("Manned Submersible", "Submersible",platforms_df$platforms)
platforms_df$fields<- gsub("Management and Conservation", "Management & Conservation",platforms_df$fields)

# Set in the right order
platforms_df$platforms <- factor(platforms_df$platforms, levels = c("SCUBA","Submersible","Rebreather","ROV","Dredging","Fishing","BRUV","AUV","Sonar"))


# Make the plot of the top fields 
# Facetwrap by platform, display only top fields
pubs_platforms_fields <- ddply(platforms_df, ~ platforms + fields,
                   function(x){c(Freq= length(x$id))})

# Measure the most important fields
pubs_fields <- ddply(platforms_df, ~ fields,
                   function(x){c(Freq_Fields= length(x$id))})
pubs_fields <- pubs_fields[order(pubs_fields$Freq_Fields, decreasing = T), ]

# Keep only the top 10 fields (this number can change)
pubs_fields_top <- pubs_fields %>% 
  arrange(desc(Freq_Fields)) %>%  slice_max(order_by = Freq_Fields, n = 12, with_ties = F)
top_fields <- unique (pubs_fields_top$fields)

# Keep in database "pubs_platforms_fields_top" the following selected fields
keep_field <-  c ("Ecology","Biodiversity","Community structure","Management & Conservation","Disturbances","Long-term monitoring","Physiology","Taxonomy","Molecular ecology", "Connectivity", "Fisheries","Geomorphology")
pubs_platforms_fields_top <- subset(pubs_platforms_fields,fields %in% keep_field)

# Set in the right order
pubs_platforms_fields_top$fields <- factor(pubs_platforms_fields_top$fields, levels = c("Geomorphology","Fisheries",  "Long-term monitoring","Management & Conservation","Disturbances","Taxonomy","Connectivity","Molecular ecology","Physiology","Ecology","Community structure","Biodiversity"))


##### Figure 3, Platforms by fields
Fig_Platforms_fields <- ggplot(pubs_platforms_fields_top, aes(x = Freq, y = fields)) + # y =   reorder (fields, Freq))
    geom_col(position = position_dodge()) + 
    facet_wrap(~platforms, labeller = label_wrap_gen(width = 18), nrow = 1) + # scales = "free_y"
    scale_x_continuous(breaks = c(0, 100,200)) +
    labs(y = "", x = "Publications") +  
    theme_bw() + theme (axis.text.x = element_text(angle = 90, hjust = 1), 
                        axis.text = element_text(size = 10), 
                        strip.text = element_text(size = 8), 
                        strip.background = element_rect(fill="white")) # ,axis.text.y = element_blank()
Fig_Platforms_fields

ggsave ( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Fig_Platforms_fields.pdf", Fig_Platforms_fields,width = 8.5, height = 3.5)
ggsave ( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Fig_Platforms_fields.jpeg", Fig_Platforms_fields,width = 8.5, height = 3.5)
ggsave ( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Fig_Platforms_fields.svg", Fig_Platforms_fields,width = 8.5, height = 3.5)

# For visual reasons and figure making, I saved the plot without yaxis
# ggsave ( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Fig_Platforms_fields_noyaxis.pdf", Fig_Platforms_fields,width = 8.5, height = 3.5)


rm (top_fields,pubs_platforms_pubs_ecosystem, pubs_platforms_depth_median, pubs_platforms_fieldspubls, order, fields, publs_complete, pubs_platforms_depth, pubs_platforms_fields_top,pubs_fields_top,pubs_fields, pubs_platforms_fields, platforms_df, Fig_Platforms_fields, pubs_platforms )

```


Make Area plot considering min and max depths. Like the figure from the previous manuscript (Bongaerts et al 2019) or website stats: http://www.mesophotic.org/statistics
Loop to account publications if it gets inside the specific depth target
Make the figure for all other groups

Supplementary Figure S2 For corals and fish
Supplementary Figure S3 For the rest of groups

- filtering publications with a original data, validated, stats and a single platform only, 

```{r}
publs <- subset(publs, publication_type == "scientific" & publication_format == "article")
publs_data <- publs[c("id","validated","included_in_stats","publication_year","original_data", "platforms", "min_depth", "max_depth", "focusgroups" )]
publs_data <- subset (publs_data, original_data == "true")
publs_data <- subset (publs_data, validated == "true")
publs_data <- subset (publs_data, included_in_stats == "true")

# Work on the Depth 
# Filter to publications having different max_depth than 0
depths_df <- filter(publs_data, max_depth != 0) 
# Or publications with max_depth in the mesophotic range
depths_df <- depths_df %>% filter(max_depth > 29) # Dropped publications where max_depth > 29
# To start making the plot from 30 m like in the website

# Transform publications where min_depth is NA to Max_Depth
# How many do we have?
sum(is.na(depths_df$min_depth))
# Replace by max depth
depths_df$min_depth <- ifelse(is.na(depths_df$min_depth), depths_df$max_depth, depths_df$min_depth)
# check
sum(is.na(depths_df$max_depth)) 
# These are the single depth publication 

# Filter publications deeper than mesophotic for the min_depth
depths_df <- depths_df %>% filter(min_depth < 151) # Keep publications where min_depth is less than 150 m
# Now we are within the mesophotic range


# Make the loop to generate a second dataframe and account for publications which min-max depths is inside the specific target depth 

# Loop should be like:
# for n from 30 to 150 depth
# Is it inside min and max, yes, account for this depth
# It is outside min and max, no, do not account for this depth

# How many pubs are between max and min depth

# 1st checking the numbers for particular depths
checkdepth = 31 # (you can check 50, 4000, etc. m)
nrow(filtered_rows <- depths_df %>%
  filter(checkdepth >= min_depth & checkdepth <= max_depth))

# Check another methodology
sum( depths_df$min_depth <= checkdepth & depths_df$max_depth >= checkdepth)
sum( checkdepth <= depths_df$max_depth & checkdepth >= depths_df$min_depth)
# They both give the same, regardless of the order


# Make the loop:
# Initialize an empty data frame to store results
result_df <- data.frame(checkdepth = numeric(0), PubNumber = numeric(0))
# Iterate through each checkdepth value
# val can be from 1 or from 30 (mesophotic)
for (val in 1:max(depths_df$max_depth)) { 
  filtered_rows <- depths_df %>%
    filter(val >= min_depth & val <= max_depth)  # Filter rows for the current value
  PubNumber <- nrow(filtered_rows)  # Count the rows
  result_df <- rbind(result_df, data.frame(checkdepth = val, PubNumber = PubNumber))
}
# Display the resulting data frame
result_df

# Let's filter the depths (from 30 to 150 m)
result_df <- subset (result_df, checkdepth<151)
result_df <- subset (result_df, checkdepth>29)


# Make the gg plot for all groups together
ggplot(result_df, aes(x = checkdepth, y = PubNumber)) +
  geom_area(fill = "cornflowerblue") +
  # scale_x_continuous(limits = c(30, 60))+ # breaks = c(30:150:10) 
  scale_x_reverse(limits = c(150, 30),breaks = seq(150,30,-10) )+
  scale_y_continuous(breaks = c(0,100,200,300,400,500,600,700))+
    coord_flip()+
    labs(title = "All focus groups",
    x = "Depths (m)",
    y = "Publications") +
  theme_bw()


# Make the figure for each group. 
# First of all, Separate all the main focusgroups in rows // Important command: 
depths_df <- separate_rows(depths_df, "focusgroups", sep=";\\s+") 

# Change names to simplify life later
depths_df$focusgroups<- gsub("Scleractinia \\(Hard Corals\\)", "Scleractinian",depths_df$focusgroups)
depths_df$focusgroups<- gsub("Octocorallia \\(Soft Corals\\)", "Octocorals",depths_df$focusgroups) # Unnecessary if onyl corals vs fish
depths_df$focusgroups<- gsub("Overall benthic \\(groups\\)", "Benthos",depths_df$focusgroups) # Unnecessary if only corals vs fish

# Check number of publications per group
ddply(depths_df, ~ focusgroups,
                   function(x){c(Pubs_number= length(x$id))})


##### Make figure plots for Fishes and Corals separately, Fig S2

### FISHES 
# Filter the data for fishes only 
depths_df_fishes <- subset(depths_df,focusgroups == "Fishes")

# Measure the number of publication where depth range is between 30 - 150 m
ddply(depths_df_fishes, ~ focusgroups,
                   function(x){c(Pubs_number= length(x$id))})

# Make the loop
# Iterate through depth, already from 30 to 150 (mesophotic)
fish_df <- data.frame(checkdepth = numeric(0), PubNumber = numeric(0))
for (val in 30:150) { # max(depths_df_fishes$max_depth)
  filtered_rows <- depths_df_fishes %>%
    filter(val >= min_depth & val <= max_depth)  # Filter rows for the current value
  PubNumber <- nrow(filtered_rows)  # Count the rows
  fish_df <- rbind(fish_df, data.frame(checkdepth = val, PubNumber = PubNumber))
}
# Display the resulting data frame
# view (fish_df)

# Make the ggplot for fishes
Fig_S3_Fish <- ggplot(fish_df, aes(x = checkdepth, y = PubNumber)) +
  geom_area(fill = "#440154FF") +
  # scale_x_continuous(limits = c(30, 60))+ # breaks = c(30:150:10) 
  scale_x_reverse(limits = c(150, 30),breaks = seq(150,30,-10) )+
  scale_y_continuous(limits = c(0,250),breaks = c(0,50,100,150,200,250))+
    coord_flip()+
    labs(title = "Fishes",
    x = "Depth (m)",
    y = "Publications") +
  theme_bw()
Fig_S3_Fish

### SCLERACTINIAN - Corals
depths_df_scleractinian <- subset(depths_df,focusgroups == "Scleractinian")

# Number of pubs sclractinian
ddply(depths_df_scleractinian, ~ focusgroups,
                   function(x){c(Pubs_number= length(x$id))})

# Make the loop
# Iterate through depth, already from 30 to 150 (mesophotic)
scleractinian_df <- data.frame(checkdepth = numeric(0), PubNumber = numeric(0))
for (val in 30:150) { 
  filtered_rows <- depths_df_scleractinian %>%
    filter(val >= min_depth & val <= max_depth)  # Filter rows for the current value
  PubNumber <- nrow(filtered_rows)  # Count the rows
  scleractinian_df <- rbind(scleractinian_df, data.frame(checkdepth = val, PubNumber = PubNumber))
}
# Display the resulting data frame
# view (scleractinian_df)

# Make the ggplot for scleractiian
Fig_S3_Scleractinia <- ggplot(scleractinian_df, aes(x = checkdepth, y = PubNumber)) +
  geom_area(fill = "#fde725FF") +
  # scale_x_continuous(limits = c(30, 60))+ # breaks = c(30:150:10) 
  scale_x_reverse(limits = c(150, 30),breaks = seq(150,30,-10) )+
  scale_y_continuous(limits = c(0,250),breaks = c(0,50,100,150,200,250))+
    coord_flip()+
    labs(title = "Scleractinian corals",
    x = "",
    y = "Publications") +
  theme_bw()
Fig_S3_Scleractinia

# Make the Figure S2 witht he combination of both
(Figure_S2 = Fig_S3_Fish + Fig_S3_Scleractinia + plot_layout(guides = 'collect', nrow = 1))

ggsave( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Fig_S2_Great_Divide_Area_plot.pdf", Figure_S2,  width = 7, height = 3)
ggsave( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Fig_S2_Great_Divide_Area_plot.jpeg", Figure_S2,  width = 7, height = 3)
ggsave( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Fig_S2_Great_Divide_Area_plot.svg", Figure_S2,  width = 7, height = 3)

#### Break in the figures ####
# Get numbers for the manuscript of the great divide for the MS
great_divide <- rbind (depths_df_fishes, depths_df_scleractinian)
length (unique(great_divide$id))
great_divide[duplicated(great_divide[,1]),]
# 614 publications which are studying either coral, fish or both
# 16 publications including both focal taxa, studying both at the same time: 152, 160, 257, 285, 306, 385, 721, 796, 799, 817, 818, 823, 966, 1046, 1424, 1528
#### Break in the figures ####


##### Make the Fig. S3 for the other groups

### Octocorals "seagreen",
depths_df_octocorals <- subset(depths_df,focusgroups == "Octocorals")

# Make the loop
# Iterate through depth, already from 30 to 150 (mesophotic)
octocorals_df <- data.frame(checkdepth = numeric(0), PubNumber = numeric(0))
for (val in 30:150) { 
  filtered_rows <- depths_df_octocorals %>%
    filter(val >= min_depth & val <= max_depth)  # Filter rows for the current value
  PubNumber <- nrow(filtered_rows)  # Count the rows
  octocorals_df <- rbind(octocorals_df, data.frame(checkdepth = val, PubNumber = PubNumber))
}
# Display the resulting data frame
# view (octocorals_df)

# Make the ggplot for octocorals
Fig_Octo <- ggplot(octocorals_df, aes(x = checkdepth, y = PubNumber)) +
  geom_area(fill = "blueviolet") +
  # scale_x_continuous(limits = c(30, 60))+ # breaks = c(30:150:10) 
  scale_x_reverse(limits = c(150, 30),breaks = seq(150,30,-10) )+
  scale_y_continuous(limits = c(0,250),breaks = c(0,50,100,150,200,250))+
    coord_flip()+
    labs(title = "Octocorals",
    x = "",
    y = "Publications") +
  theme_bw()+ theme (plot.title = element_text(size = 11))
Fig_Octo


### Benthos "gray62",
depths_df_benthos <- subset(depths_df,focusgroups == "Benthos")

# Make the loop
# Iterate through depth, already from 30 to 150 (mesophotic)
benthos_df <- data.frame(checkdepth = numeric(0), PubNumber = numeric(0))
for (val in 30:150) { 
  filtered_rows <- depths_df_benthos %>%
    filter(val >= min_depth & val <= max_depth)  # Filter rows for the current value
  PubNumber <- nrow(filtered_rows)  # Count the rows
  benthos_df <- rbind(benthos_df, data.frame(checkdepth = val, PubNumber = PubNumber))
}
# Display the resulting data frame
# view (benthos_df)

# Make the ggplot for benthos
Fig_Benthos <- ggplot(benthos_df, aes(x = checkdepth, y = PubNumber)) +
  geom_area(fill = "gray62") +
  # scale_x_continuous(limits = c(30, 60))+ # breaks = c(30:150:10) 
  scale_x_reverse(limits = c(150, 30),breaks = seq(150,30,-10) )+
  scale_y_continuous(limits = c(0,250),breaks = c(0,50,100,150,200,250))+
    coord_flip()+
    labs(title = "Overall benthic groups",
    x = "Depths (m)",
    y = "Publications") +
  theme_bw()+ theme (plot.title = element_text(size = 11))
Fig_Benthos


### Algae (Macro, Turf and Crustose Coralline)
depths_df_algae <- subset(depths_df,focusgroups == "Algae (Macro, Turf and Crustose Coralline)")

# Make the loop
# Iterate through depth, already from 30 to 150 (mesophotic)
algae_df <- data.frame(checkdepth = numeric(0), PubNumber = numeric(0))
for (val in 30:150) { 
  filtered_rows <- depths_df_algae %>%
    filter(val >= min_depth & val <= max_depth)  # Filter rows for the current value
  PubNumber <- nrow(filtered_rows)  # Count the rows
  algae_df <- rbind(algae_df, data.frame(checkdepth = val, PubNumber = PubNumber))
}
# Display the resulting data frame
# view (blackcorals_df)

# Make the ggplot for benthos
Fig_Algae <- ggplot(algae_df, aes(x = checkdepth, y = PubNumber)) +
  geom_area(fill = "green") +
  # scale_x_continuous(limits = c(30, 60))+ # breaks = c(30:150:10) 
  scale_x_reverse(limits = c(150, 30),breaks = seq(150,30,-10) )+
  scale_y_continuous(limits = c(0,250),breaks = c(0,50,100,150,200,250))+
    coord_flip()+
    labs(title = "Algae (Macro, Turf and CCA)",
    x = "",
    y = "Publications") +
  theme_bw()+ theme (plot.title = element_text(size = 11))
Fig_Algae


### Porifera (Sponges)
depths_df_sponges <- subset(depths_df,focusgroups == "Porifera (Sponges)")

# Make the loop
# Iterate through depth, already from 30 to 150 (mesophotic)
sponges_df <- data.frame(checkdepth = numeric(0), PubNumber = numeric(0))
for (val in 30:150) { 
  filtered_rows <- depths_df_sponges %>%
    filter(val >= min_depth & val <= max_depth)  # Filter rows for the current value
  PubNumber <- nrow(filtered_rows)  # Count the rows
  sponges_df <- rbind(sponges_df, data.frame(checkdepth = val, PubNumber = PubNumber))
}
# Display the resulting data frame
# view (sponges_df)

# Make the ggplot for benthos
Fig_Sponges <- ggplot(sponges_df, aes(x = checkdepth, y = PubNumber)) +
  geom_area(fill = "coral1") +
  # scale_x_continuous(limits = c(30, 60))+ # breaks = c(30:150:10) 
  scale_x_reverse(limits = c(150, 30),breaks = seq(150,30,-10) )+
  scale_y_continuous(limits = c(0,250),breaks = c(0,50,100,150,200,250))+
    coord_flip()+
    labs(title = "Porifera (Sponges)",
    x = "",
    y = "Publications") +
  theme_bw()+ theme (plot.title = element_text(size = 11))
Fig_Sponges


### Antipatharia (Black Corals)
depths_df_bc <- subset(depths_df,focusgroups == "Antipatharia (Black Corals)")

# Make the loop
# Iterate through depth, already from 30 to 150 (mesophotic)
blackcorals_df <- data.frame(checkdepth = numeric(0), PubNumber = numeric(0))
for (val in 30:150) { 
  filtered_rows <- depths_df_bc %>%
    filter(val >= min_depth & val <= max_depth)  # Filter rows for the current value
  PubNumber <- nrow(filtered_rows)  # Count the rows
  blackcorals_df <- rbind(blackcorals_df, data.frame(checkdepth = val, PubNumber = PubNumber))
}
# Display the resulting data frame
# view (blackcorals_df)

# Make the ggplot for benthos
Fig_Blackcorals <- ggplot(blackcorals_df, aes(x = checkdepth, y = PubNumber)) +
  geom_area(fill = "black") +
  # scale_x_continuous(limits = c(30, 60))+ # breaks = c(30:150:10) 
  scale_x_reverse(limits = c(150, 30),breaks = seq(150,30,-10) )+
  scale_y_continuous(limits = c(0,250),breaks = c(0,50,100,150,200,250))+
    coord_flip()+
    labs(title = "Antipatharia (Black Corals)",
    x = "Depths (m)",
    y = "Publications") +
  theme_bw()+ theme (plot.title = element_text(size = 11))
Fig_Blackcorals


### Crustacea
depths_df_crustacea <- subset(depths_df,focusgroups == "Crustacea")

# Make the loop
# Iterate through depth, already from 30 to 150 (mesophotic)
crustacea_df <- data.frame(checkdepth = numeric(0), PubNumber = numeric(0))
for (val in 30:150) { 
  filtered_rows <- depths_df_crustacea %>%
    filter(val >= min_depth & val <= max_depth)  # Filter rows for the current value
  PubNumber <- nrow(filtered_rows)  # Count the rows
  crustacea_df <- rbind(crustacea_df, data.frame(checkdepth = val, PubNumber = PubNumber))
}
# Display the resulting data frame
# view (crustacea_df)

# Make the ggplot for benthos
Fig_Crustacea <- ggplot(crustacea_df, aes(x = checkdepth, y = PubNumber)) +
  geom_area(fill = "brown3") +
  # scale_x_continuous(limits = c(30, 60))+ # breaks = c(30:150:10) 
  scale_x_reverse(limits = c(150, 30),breaks = seq(150,30,-10) )+
  scale_y_continuous(limits = c(0,250),breaks = c(0,50,100,150,200,250))+
    coord_flip()+
    labs(title = "Crustacea",
    x = "",
    y = "Publications") +
  theme_bw()+ theme (plot.title = element_text(size = 11))
Fig_Crustacea


### Other invertebrates
depths_df_otherinvertebrates <- subset(depths_df,focusgroups == "Other invertebrates")

# Make the loop
# Iterate through depth, already from 30 to 150 (mesophotic)
inv_df <- data.frame(checkdepth = numeric(0), PubNumber = numeric(0))
for (val in 30:150) { 
  filtered_rows <- depths_df_otherinvertebrates %>%
    filter(val >= min_depth & val <= max_depth)  # Filter rows for the current value
  PubNumber <- nrow(filtered_rows)  # Count the rows
  inv_df <- rbind(inv_df, data.frame(checkdepth = val, PubNumber = PubNumber))
}
# Display the resulting data frame
# view (inv_df)

# Make the ggplot for benthos
Fig_Other_inv <- ggplot(inv_df, aes(x = checkdepth, y = PubNumber)) +
  geom_area(fill = "darkgoldenrod1") +
  # scale_x_continuous(limits = c(30, 60))+ # breaks = c(30:150:10) 
  scale_x_reverse(limits = c(150, 30),breaks = seq(150,30,-10) )+
  scale_y_continuous(limits = c(0,250),breaks = c(0,50,100,150,200,250))+
    coord_flip()+
    labs(title = "Other invertebrates",
    x = "",
    y = "Publications") +
  theme_bw()+ theme (plot.title = element_text(size = 11))
Fig_Other_inv


### Symbiodinium (zooxanthellae)
depths_df_Symbio <- subset(depths_df,focusgroups == "Symbiodinium (zooxanthellae)")

# Make the loop
# Iterate through depth, already from 30 to 150 (mesophotic)
symbio_df <- data.frame(checkdepth = numeric(0), PubNumber = numeric(0))
for (val in 30:150) { 
  filtered_rows <- depths_df_Symbio %>%
    filter(val >= min_depth & val <= max_depth)  # Filter rows for the current value
  PubNumber <- nrow(filtered_rows)  # Count the rows
  symbio_df <- rbind(symbio_df, data.frame(checkdepth = val, PubNumber = PubNumber))
}
# Display the resulting data frame
# view (symbio_df)

# Make the ggplot for benthos
Fig_Symbio <- ggplot(symbio_df, aes(x = checkdepth, y = PubNumber)) +
  geom_area(fill = "darkolivegreen1") +
  # scale_x_continuous(limits = c(30, 60))+ # breaks = c(30:150:10) 
  scale_x_reverse(limits = c(150, 30),breaks = seq(150,30,-10) )+
  scale_y_continuous(limits = c(0,250),breaks = c(0,50,100,150,200,250))+
    coord_flip()+
    labs(title = "Symbiodinium (zooxanthellae)",
    x = "",
    y = "Publications") +
  theme_bw()+ theme (plot.title = element_text(size = 11))
Fig_Symbio

#### Make Fig S3 combining them all

(Figure_S3 = Fig_Benthos + Fig_Algae + Fig_Sponges + Fig_Octo + Fig_Blackcorals + Fig_Symbio + Fig_Crustacea + Fig_Other_inv +plot_layout(guides = 'collect', nrow = 2))

ggsave( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Figure_S3.pdf", Figure_S3,  width = 11, height = 6)
ggsave( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Figure_S3.jpeg", Figure_S3,  width = 11, height = 6)
ggsave( "~/Documents/AAASea_Science/Mesophotic.org/Figures/Figure_S3.svg", Figure_S3,  width = 11, height = 6)

rm (depths_df, benthos_df, octocorals_df, algae_df, blackcorals_df, crustacea_df, depths_df_algae,depths_df_bc,depths_df_benthos,depths_df_crustacea,depths_df_fishes, depths_df_octocorals,depths_df_otherinvertebrates,depths_df_scleractinian,depths_df_sponges,depths_df_Symbio,Fig_Algae,Fig_Benthos,Fig_Blackcorals,Fig_Crustacea,Fig_Octo,Fig_Other_inv,Fig_S3_Fish,Fig_S3_Scleractinia,Fig_Sponges,Fig_Symbio,Figure_S3,fish_df,great_divide,filtered_rows,inv_df,scleractinian_df,sponges_df,symbio_df,result_df, Figure_S2)

```


## THE END OF FIGURES USED IN THE MANUSCRIPT

From here below is additional. Codes are not necessarily clean. 

Plot of platforms frequency 
```{r}
publs <- subset(publs, publication_type == "scientific" & publication_format == "article")
publs_data <- publs[c("id","validated","included_in_stats","publication_year","original_data","fields","focusgroups", "platforms", "locations")]

publs_data <- subset (publs_data, original_data == "true")
publs_data <- subset (publs_data, validated == "true")
publs_data <- subset (publs_data, included_in_stats == "true")

# Split `Platforms` into new rows, but keep ID
publs_platforms <- separate_rows(publs_data, "platforms", sep=";\\s+") 

# Remove rows with no associated `platforms`
publs_platforms <- publs_platforms[!
(is.na(publs_platforms$platforms) |
publs_platforms$platforms ==
""), ]

platforms <- as.data.frame(table(publs_platforms$platforms))
colnames (platforms) <- c("Platform", "freq")

# Measure in respect to different to mean
platforms$media_freq <- platforms$freq - median (platforms$freq) # or mean 

platforms <- arrange(platforms, media_freq)
platforms$Platform <- factor(platforms$Platform, levels = platforms$Platform)
# Make the plot

ggplot(platforms, aes(Platform, media_freq, fill = Platform)) + 
  geom_col(position = position_dodge()) + 
  labs(y = "Frequency based on median", x = "", 
  title = "Platform") + coord_flip() + 
  theme_bw() + theme(legend.position="none")

rm (platforms, publs_data, publs_platforms)

```

Another plot by platforms 
```{r}

publs <- subset(publs, publication_type == "scientific" & publication_format == "article")

publs_data <- publs[c("id","validated","included_in_stats","publication_year","original_data","fields","focusgroups", "platforms", "locations")]

publs_data <- subset (publs_data, original_data == "true")
publs_data <- subset (publs_data, validated == "true")
publs_data <- subset (publs_data, included_in_stats == "true")

# Number of publications per year
publs_data1 <- ddply(publs_data, ~ publication_year,
                   function(x){c(NumbPubs= length(x$id)) })

publs_data2 <- dplyr::filter(publs_data, grepl('Rebreather', platforms)) # You can add extra categories like |Remotely Operated Vehicle (ROV)|Manned Submersible...

publs_data3 <- dplyr::filter(publs_data, grepl('Manned Submersible', platforms)) 

publs_data4 <- dplyr::filter(publs_data, grepl('SCUBA', platforms)) 

## Number of publications using Rebreather per year!
pubs_rebreather <- ddply(publs_data2, ~ publication_year,
                   function(x){c(pubs_rebreather= length(x$id))})

## Number of publications using Submersible per year!
pubs_Submersible <- ddply(publs_data3, ~ publication_year,
                   function(x){c(pubs_Submersible= length(x$id))})

## Number of publications using SCUBA!
pubs_SCUBA <- ddply(publs_data4, ~ publication_year,
                   function(x){c(pubs_SCUBA= length(x$id))})


# Merge all dataframes!
pubs_human_access <- merge (publs_data1, pubs_rebreather, by.x = "publication_year", all = T )
pubs_human_access <- merge (pubs_human_access, pubs_Submersible, by.x = "publication_year", all = T )
pubs_human_access <- merge (pubs_human_access, pubs_SCUBA, by.x = "publication_year", all = T )

# Replace all NA by 0 
pubs_human_access <- pubs_human_access %>% replace(is.na(.), 0)

# Cummulative Rolling sum 
pubs_human_access$All_Pub <- cumsum(pubs_human_access$NumbPubs)
pubs_human_access$Reb_Pub <- cumsum(pubs_human_access$pubs_rebreather)
pubs_human_access$Sub_Pub <- cumsum(pubs_human_access$pubs_Submersible)
pubs_human_access$SCUBA_Pub <- cumsum(pubs_human_access$pubs_SCUBA)
# They are defined as "_Pub"

# Accumulative representation
pubs_human_access_acumm <- pubs_human_access[c("publication_year","All_Pub","Reb_Pub","Sub_Pub","SCUBA_Pub")]
df_accum <- melt(pubs_human_access_acumm ,  id.vars = 'publication_year', variable.name = 'series')
# Make the plot 
ggplot(df_accum, aes(publication_year, value)) +
  geom_line(aes(colour = series)) + labs(y = "Number Publications", x= "Years", 
                                         title = "Human access to Mesophotic depths") +
  theme_bw()

# Non-Accummulative
pubs_human_access_non_accumm <- pubs_human_access[c("publication_year","NumbPubs","pubs_rebreather","pubs_Submersible","pubs_SCUBA")]
df_non_accumm <- melt(pubs_human_access_non_accumm ,  id.vars = 'publication_year', variable.name = 'series')
# Make the plot
ggplot(df_non_accumm, aes(publication_year, value)) +
  geom_line(aes(colour = series)) + labs(y = "Number Publications", x= "Years", 
  title = "Human access to Mesophotic depths")  +
  theme_bw()


rm (publs_data1, publs_data2,publs_data3,publs_data4, pubs_rebreather, pubs_SCUBA, pubs_Submersible, pubs_human_access, pubs_human_access_acumm,pubs_human_access_non_accumm, pubs_human_access, df_accum, df_non_accumm)

```

Plot by depths. Publications and focus groups. Max_depth target vs depth ranges target

```{r}

publs <- subset(publs, publication_type == "scientific" & publication_format == "article")
publs_data <- publs[c("id","validated","included_in_stats","publication_year","original_data", "platforms", "min_depth", "max_depth", "focusgroups" )]
publs_data <- subset (publs_data, original_data == "true")
publs_data <- subset (publs_data, validated == "true")
publs_data <- subset (publs_data, included_in_stats == "true")

# Delete columns where platforms is empty
depths_df <- publs_data[-which(publs_data$platforms == ""), ]

# Filter to publications having different max_depth than 0
depths_df <- filter(depths_df, max_depth != 0)
# Or publications with max_depth in the mesophotic range
depths_df <- depths_df %>% filter(max_depth > 29) # Dropped publications where max_depth > 29


# Transform publications where min_depth is NA to Max_Depth
# How many do we have?
sum(is.na(depths_df$min_depth))
# Replace by max depth
depths_df$min_depth <- ifelse(is.na(depths_df$min_depth), depths_df$max_depth, depths_df$min_depth)
# check
sum(is.na(depths_df$max_depth)) # check the max depths

# Filter publications deeper than mesophotic for the min_depth
depths_df <- depths_df %>% filter(min_depth < 151) # Keep publications where min_depth is less than 150 m


# Transform all rows where "max_depth" is more than 300 (as outside of mesophotic range)
# Should be unnecessary based on plot after
depths_df$max_depth <- ifelse(depths_df$max_depth>300,300,depths_df$max_depth)

# Measure publications depth range
depths_df$Depth_range <- depths_df$max_depth - depths_df$min_depth

# THIS SHOULD BE UNNECESSARY based on command above- Transform all values where "Depth_range" is more than 300 (as outside of mesophotic range) 
depths_df$Depth_range <- ifelse(depths_df$Depth_range>300,300,depths_df$Depth_range)

# Quick check histograms
hist (depths_df$Depth_range)


# scale the variable, target max depths and not depth ranges
depths_df$max_depth_target = cut(depths_df$max_depth,c(29,40,60,80,100,150,300)) # 3962
unique (depths_df$max_depth_target)
# You can also choose the number: , breaks = 5
levels(depths_df$max_depth_target) = c("30-40 m","40-60 m", "60-80 m", "80-100 m", "100-150 m", "> 150 m")

# This would be for Group_Depth_ranges
depths_df$group_depth_range = cut(depths_df$Depth_range,c(-1,1,10,50,100,300)) # 3962
unique (depths_df$group_depth_range)
# You can also choose the number: , breaks = 5
levels(depths_df$group_depth_range) = c("0 m","1-10 m", "10-50 m", "50-100 m", "> 100 m")

ggplot(data =depths_df,aes(x=max_depth_target)) + 
  geom_bar() +
  labs(x="Depth range (m)", y="Number of publications") + theme_bw()


#  set factors in the right order
depths_df$max_depth_target <- factor(depths_df$max_depth_target, levels = c("> 150 m","100-150 m","80-100 m","60-80 m","40-60 m", "30-40 m"))

ggplot(data =depths_df,aes(y=max_depth_target)) +
  geom_bar(position = position_dodge()) +
  labs(x="Number of publications", y="Maximum depths (m)") + 
  scale_x_continuous(position = "bottom", breaks = c(0, 50, 100,150, 200,250))+
  theme_bw() + 
  theme (legend.position="none", legend.title=element_blank())


# Quick check: 
ddply(depths_df, ~ max_depth_target,
                   function(x){c(Pubs_number_all= length(x$id))})

# Important command: Separate all the main focusgroups
depths_df <- separate_rows(depths_df, "focusgroups", sep=";\\s+") 

# Only the great divide, corals vs fish
keep_focus_groups <- c("Fishes", "Scleractinia (Hard Corals)") 

# Another option for keeping other focus groups
# keep_focus_groups <- c("Fishes", "Scleractinia (Hard Corals)","Octocorallia (Soft Corals)", "Overall benthic (groups)") # "Algae (Macro, Turf and Crustose Coralline)", "Porifera (Sponges)"

depths_df <- subset(depths_df,focusgroups %in% keep_focus_groups)


# Change names to simplify life
depths_df$focusgroups<- gsub("Scleractinia \\(Hard Corals\\)", "Scleractinian",depths_df$focusgroups)
# Unnecessary if only corals vs fish
# depths_df$focusgroups<- gsub("Octocorallia \\(Soft Corals\\)", "Octocorals",depths_df$focusgroups) 
# depths_df$focusgroups<- gsub("Overall benthic \\(groups\\)", "Benthos",depths_df$focusgroups) # Unnecessary if onyl corals vs fish

# Check number of publications per focus groups
ddply(depths_df, ~ focusgroups,
                   function(x){c(Pubs_number= length(x$id))})
# Numbers are a bit lower than the mentioned in the manuscript because we are droping the publications where platform is empty

# Set colours manually, order and colour
depths_df$focusgroups <- factor(depths_df$focusgroups, levels = c("Fishes", "Scleractinian")) # Only Fishes and Scleractinian
viridisPalette5 <- c("#440154FF", "#fde725FF")

# depths_df$focusgroups <- factor(depths_df$focusgroups, levels = c("Fishes", "Octocorals", "Scleractinian", "Benthos")) # Unnecessary if only corals vs fish
# viridisPalette5 <- c("#440154FF", "#414487FF", "#fde725FF", "gray62")

ggplot(data =depths_df,aes(x=group_depth_range, fill = focusgroups)) + 
  geom_bar(position = position_dodge()) +
  labs(x="Depth range (m)", y="Number of publications") + 
  scale_fill_manual(values=viridisPalette5) + theme_bw() + 
  theme (legend.position="bottom", legend.title=element_blank())

# 2nd option
ggplot(data =depths_df,aes(x=group_depth_range, fill = focusgroups)) + 
  geom_bar(position = position_dodge()) +
  facet_wrap(~focusgroups) +
  labs(x="Depth range (m)", y="Number of publications") + 
  scale_fill_manual(values=viridisPalette5) + theme_bw() + 
  theme (legend.position="none", legend.title=element_blank())


# Instead of depth ranges, target max_depths by groups!
# scale the variable max_depth
depths_df$max_depth_target = cut(depths_df$max_depth,c(29,40,60,80,100,150,300)) # 3962
unique (depths_df$max_depth_target)
# You can also choose the number: , breaks = 5
levels(depths_df$max_depth_target) = c("30-40 m","40-60 m", "60-80 m", "80-100 m", "100-150 m", "> 150 m")

# Make a manual ddply to verify the values
summary_groups_depths <- ddply(depths_df, ~ focusgroups + max_depth_target,
                   function(x){c(Pubs_number= length(x$id))})

ggplot(data =depths_df,aes(x=max_depth_target, fill = focusgroups)) + 
  geom_bar(position = position_dodge()) +
  labs(x="Max depth target (m)", y="Number of publications") + 
  scale_fill_manual(values=viridisPalette5) + theme_bw() + 
  theme (legend.position="bottom", legend.title=element_blank())


# 2nd option with all 4
ggplot(data =depths_df,aes(x=max_depth_target, fill = focusgroups)) + 
  geom_bar(position = position_dodge()) +
  facet_wrap(~focusgroups, nrow = 1) +
  labs(x="Max depth target (m)", y="Number of publications") + 
  scale_fill_manual(values=viridisPalette5) + theme_bw() + 
  theme (legend.position="none", legend.title=element_blank(),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


# Necessary to set the order of y axis manually again
depths_df$max_depth_target <- factor(depths_df$max_depth_target, levels = c("> 150 m","100-150 m","80-100 m","60-80 m","40-60 m", "30-40 m"))

ggplot(data =depths_df,aes(y=max_depth_target, fill = focusgroups)) +
  geom_bar(position = position_dodge()) +
  facet_wrap(~focusgroups, nrow = 1) +
  labs(x="Number of publications", y="Maximum depths (m)") + 
  scale_x_continuous(position = "bottom", breaks = c(0, 25, 50,75, 100))+
  scale_fill_manual(values=viridisPalette5) + theme_bw() + 
  theme (legend.position="none", legend.title=element_blank())


# Check of figure with new summary dataframe
summary_groups_depths$max_depth_target <- factor(summary_groups_depths$max_depth_target, levels = c("> 150 m","100-150 m","80-100 m","60-80 m","40-60 m", "30-40 m"))

ggplot(data =summary_groups_depths,aes(y=max_depth_target,x=Pubs_number, fill = focusgroups)) +
  geom_bar(stat='identity') +
  facet_wrap(~focusgroups, nrow = 1) +
  labs(x="Number of publications", y="Maximum depths (m)") + 
  scale_x_continuous(position = "bottom", breaks = c(0, 25, 50,75, 100))+
  scale_fill_manual(values=viridisPalette5) + theme_bw() + 
  theme (legend.position="none", legend.title=element_blank())

rm (pubs_platforms, summary_groups_depths, platforms_df, depths_df)

```


Check number of publications per focus groups
For long-term

```{r}
publs <- subset(publs, publication_type == "scientific" & publication_format == "article")
publs_data <- publs[c("id","validated","included_in_stats","publication_year","original_data","min_depth", "max_depth", "focusgroups", "fields")]
publs_data <- subset (publs_data, original_data == "true")
publs_data <- subset (publs_data, validated == "true")
publs_data <- subset (publs_data, included_in_stats == "true")

# Filter to publications having different max_depth than 0
depths_df <- filter(publs_data, max_depth != 0) # If you run command above, change "publs_data" with "depths_df"
# Or publications with max_depth in the mesophotic range
depths_df <- depths_df %>% filter(max_depth > 29) # Dropped publications where max_depth > 29
# To start making the plot from 30 m like in the website

# Transform publications where min_depth is NA to Max_Depth
# How many do we have?
sum(is.na(depths_df$min_depth))
# Replace by max depth
depths_df$min_depth <- ifelse(is.na(depths_df$min_depth), depths_df$max_depth, depths_df$min_depth)
# check
sum(is.na(depths_df$max_depth)) 

# Filter publications deeper than mesophotic for the min_depth
depths_df <- depths_df %>% filter(min_depth < 151) # Keep publications where min_depth is less than 150 m


# Making it for Fishes and corals separately
# First of all, Separate all the main focusgroups in rows // Important command: 
depths_df <- separate_rows(depths_df, "focusgroups", sep=";\\s+") 

# Change names to simplify life later
depths_df$focusgroups<- gsub("Scleractinia \\(Hard Corals\\)", "Scleractinian",depths_df$focusgroups)
depths_df$focusgroups<- gsub("Octocorallia \\(Soft Corals\\)", "Octocorals",depths_df$focusgroups) # Unnecessary if onyl corals vs fish
depths_df$focusgroups<- gsub("Overall benthic \\(groups\\)", "Benthos",depths_df$focusgroups) # Unnecessary if onyl corals vs fish

# Check number of publications of fishes and corals
ddply(depths_df, ~ focusgroups,
                   function(x){c(Pubs_number= length(x$id))})

depths_df <- separate_rows(depths_df, "fields", sep=";\\s+") 

long_term <- filter(depths_df, fields == "Long-term monitoring") 

ddply(long_term, ~ focusgroups,
                   function(x){c(Pubs_number= length(x$id))})

# Put scleractinian, symbiodinium, benthos, octocorals together
keep_groups <- c("Scleractinian", "Symbiodinium (zooxanthellae)", "Octocorals", "Benthos")

long_term_2 <-  long_term[long_term$focusgroups %in% keep_groups, ]

length (unique(long_term_2$id))
long_term_2[duplicated(long_term_2[,1]),] # 15 multigroup manuscripts

rm (depths_df,long_term,long_term_2)

```

Figure by platforms and locations, 
Keep locations in the same order
Info of TME and MCEs is unnecessary according to coregroup

```{r}

publs <- subset(publs, publication_type == "scientific" & publication_format == "article")
publs_data <- publs[c("id","validated","included_in_stats","publication_year","original_data","mce", "platforms","min_depth", "max_depth", "locations")]
publs_data <- subset (publs_data, original_data == "true")
publs_data <- subset (publs_data, validated == "true")
publs_data <- subset (publs_data, included_in_stats == "true")

# Replace mce column by Ecosystem
colnames(publs_data)[colnames(publs_data) == 'mce'] <- 'Ecosystem'

# Set info of MCE or TME
publs_data$Ecosystem <- ifelse(publs_data$Ecosystem %in% c('true'), "MCE","TME")

# Generate a new database
platforms_df <- publs_data

# Keep all platforms! Do not filter to publications with single platform! (if necessary check above)
# Work on PLATFORMS
# First delete columns where platforms is empty
platforms_df <- platforms_df[-which(platforms_df$platforms == ""), ]

# Work on LOCATIONS
# Remove rows with no associated `locations`
platforms_df <- platforms_df[!(is.na(platforms_df$locations) | platforms_df$locations == ""), ]

# Remove publications where max_depth == 0
platforms_df <- platforms_df %>% filter(max_depth > 29) # Dropped publications where max_depth > 29


# Transform publications where min_depth is NA to Max_Depth
# How many do we have?
sum(is.na(platforms_df$min_depth))
# Replace by max depth
platforms_df$min_depth <- ifelse(is.na(platforms_df$min_depth), platforms_df$max_depth, platforms_df$min_depth)
# check
sum(is.na(platforms_df$max_depth)) 

# Filter publications deeper than mesophotic for the min_depth
platforms_df <- platforms_df %>% filter(min_depth < 151) # Keep publications where min_depth is less than 150 m


#### BE CAREFUL - THESE COMMANDS GENERATES MORE PUBLICATIONS NUMBERS due multiplatform and multilocation - Keep all of them but as separate rows (option to be decided) #####
platforms_df <- separate_rows(platforms_df, "platforms", sep=";\\s+") 
platforms_df <- separate_rows(platforms_df, "locations", sep=";\\s+") 

# Remove string after "-" to keep only the country.
# platforms_df$locations <- sub("\\ -.*", "", platforms_df$locations)

# Delete duplicated rows because multiple locations withing the same country. 
platforms_df <- platforms_df[!duplicated(platforms_df), ]

# Check number of publications
top_locations <- ddply(platforms_df, ~ Ecosystem + locations,
                   function(x){c(Pubs_number= length(x$id))})

# Check number of publications
top_locations <- ddply(platforms_df, ~ locations,
                   function(x){c(Pubs_number= length(x$id))})

order (top_locations)

# Keep only certain platforms and arrange by depth
keep_plat <-  c ("SCUBA (open-circuit or unspecified)","Autonomous Underwater Vehicle (AUV)","Rebreather","Baited Remote Underwater Video (BRUV)","Fishing","Remotely Operated Vehicle (ROV)","Manned Submersible", "Dredging / trawling", "Sonar / Multibeam")
platforms_df <- subset(platforms_df,platforms %in% keep_plat)

# Transform characters to simplified names
platforms_df$platforms<- gsub("SCUBA \\(open-circuit or unspecified\\)", "SCUBA",platforms_df$platforms)
platforms_df$platforms<- gsub("Autonomous Underwater Vehicle \\(AUV\\)", "AUV",platforms_df$platforms)
platforms_df$platforms<- gsub("Sonar / Multibeam", "Sonar",platforms_df$platforms)
platforms_df$platforms<- gsub("Baited Remote Underwater Video \\(BRUV\\)", "BRUV",platforms_df$platforms)
platforms_df$platforms<- gsub("Dredging / trawling", "Dredging",platforms_df$platforms)
platforms_df$platforms<- gsub("Remotely Operated Vehicle \\(ROV\\)", "ROV",platforms_df$platforms)
platforms_df$platforms<- gsub("Manned Submersible", "Submersible",platforms_df$platforms)

# Set in the right order
platforms_df$platforms <- factor(platforms_df$platforms, levels = c("SCUBA","Submersible","Rebreather","ROV","Dredging","Fishing","BRUV","AUV","Sonar"))

# Make a table with the number of publications per category: 
pubs_platforms <- ddply(platforms_df, ~ platforms,
                   function(x){c(Pubs_number= length(x$id))})
pubs_platforms_depth <- ddply(platforms_df, ~ platforms,
                   function(x){c(Max_depth_mean= mean(x$max_depth))})
pubs_platforms_depth_median <- ddply(platforms_df, ~ platforms,
                   function(x){c(Max_depth_median= median(x$max_depth))})
pubs_platforms_pubs_ecosystem <- ddply(platforms_df, ~ platforms + Ecosystem,
                   function(x){c(Pubs_number_ecos= length(x$id))})
pubs_platforms <- merge (pubs_platforms,pubs_platforms_depth)
pubs_platforms <- merge (pubs_platforms,pubs_platforms_depth_median)
pubs_platforms <- merge (pubs_platforms,pubs_platforms_pubs_ecosystem)

# Make the plot of the top locations 
# Facetwrap by platform, display only top locations
pubs_platforms_locations <- ddply(platforms_df, ~ platforms + locations,
                   function(x){c(Freq= length(x$id))})

# Measure most important location
pubs_locations <- ddply(platforms_df, ~ locations,
                   function(x){c(Freq_Loc= length(x$id))})
pubs_locations <- pubs_locations[order(pubs_locations$Freq_Loc, decreasing = T), ]


# Keep only the top 10 locations (this number can change)
pubs_locations_top <- pubs_locations %>% 
  arrange(desc(Freq_Loc)) %>%  slice_max(order_by = Freq_Loc, n = 10, with_ties = F)

top_locations <- unique (pubs_locations_top$locations)

# Keep in database "pubs_platforms_locations_top" only the top 10 locations
pubs_platforms_locations_top <- subset(pubs_platforms_locations,locations %in% top_locations)

# Set in the same order
pubs_platforms_locations_top$locations <- factor(pubs_platforms_locations_top$locations, levels = top_locations)


# Fig_Platforms_Locations
ggplot(pubs_platforms_locations_top, aes(x = Freq, y = reorder (locations, Freq))) + 
  geom_col(position = position_dodge()) + 
  facet_wrap(~platforms, labeller = label_wrap_gen(width = 18), nrow = 1) + # scales = "free_y"
  scale_x_continuous(breaks = c(0, 25, 50, 75, 100)) +
  labs(y = "", x = "Number of publications (n)", title = "Top 10 locations") +  
  theme_bw() 

rm (top_locations,pubs_platforms_pubs_ecosystem, pubs_platforms_depth_median, pubs_platforms_locationspubls, order, Locations, publs_complete, pubs_platforms_depth, pubs_platforms_locations_top,pubs_locations_top,pubs_locations, Fig_Platforms_Locations,pubs_platforms_locations, pubs_platforms_depth, pubs_platforms, platforms_df)

```

Plot by focus groups and top locations

```{r}

publs <- subset(publs, publication_type == "scientific" & publication_format == "article")
publs_data <- publs[c("id","publication_year","original_data","fields","focusgroups", "platforms", "locations")]
publs_data <- subset (publs_data, original_data == "true")

# Split `focusgroups` into new rows, but keep ID
publs_focusgroups <- separate_rows(publs_data, "focusgroups", sep=";\\s+") 

# Remove rows with no associated `platforms`
publs_focusgroups <- publs_focusgroups[!
(is.na(publs_focusgroups$focusgroups) |
publs_focusgroups$focusgroups ==
""), ]

Groups <- as.data.frame(table(publs_focusgroups$focusgroups))
colnames (Groups) <- c("Focus_groups", "freq")

Groups$media_freq <- Groups$freq - median (Groups$freq)

Groups <- arrange(Groups, media_freq)
# Groups <- rbind(head(Groups,5), tail(Groups,5))
Groups$Focus_groups <- factor(Groups$Focus_groups, levels = Groups$Focus_groups)
# Make the plot

ggplot(Groups, aes(Focus_groups, media_freq, fill = Focus_groups)) + 
  geom_col(position = position_dodge()) + 
  labs(y = "Frequency based on the median", x = "", 
  title = "Focus groups") + coord_flip() + 
  theme_bw() + theme(legend.position="none")

rm (Groups, publs_focusgroups)

```

Plot by locations, filtering to top locations only

```{r}

# Split `locations` into new rows, but keep ID
publs_locations <- separate_rows(publs_data, "locations", sep=";\\s+") 

# Remove rows with no associated `platforms`
publs_locations <- publs_locations[!
(is.na(publs_locations$locations) |
publs_locations$locations ==
""), ]

Locations <- as.data.frame(table(publs_locations$locations))
colnames (Locations) <- c("Locations", "freq")

Locations$media_freq <- Locations$freq - median (Locations$freq)

Locations <- arrange(Locations, media_freq)
# Locations <- rbind(head(Locations,7), tail(Locations,7))
Locations <- rbind(tail(Locations,15))

Locations$Locations <- factor(Locations$Locations, levels = Locations$Locations)
# Make the plot

ggplot(Locations, aes(Locations, media_freq, fill = Locations)) + 
  geom_col(position = position_dodge()) + 
  labs(y = "Frequency based on the median", x = "", 
  title = "Locations") + coord_flip() + 
  theme_bw() + theme(legend.position="none")

rm (Locations, publs_locations)

```

#### THE END


