---
title: "Mesophotic_Database_release_2.0"
author: "Veronica Radice"
date: "10/18/2023"
output: html_document
---

Code for the manuscript:

*"Recent trends and biases in mesophotic ecosystem research"*

Veronica Z. Radice, Alejandra Hernandez-Agreda, Gonzalo Pérez-Rosales, Ryan Booker, Jessica Bellworthy, Manon Broadribb, Gaby E. Carpenter, Clara Diaz, Ryan J. Eckert, Nicola L. Foster, Johanna C. Gijsbers, Erika Gress, Jack H Laverick, Valerio Micaroni, Miriam Pierotti, Héloïse Rouzé, Angela Stevenson, Alexis B. Sturm, Pim Bongaerts

< insert DOI and citation >


### Load packages
```{r message=FALSE}
require(dplyr)
require(plyr)
require(tidyr)
require(ggplot2)
#require(ggstance)
require(ggpubr)
require(tidyverse)
require(cowplot)
library(RcppRoll)
library(ggridges)
```

##### Colorblind-friendly palette
```{r}
# http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
#cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

library(viridis)

viridisPalette5 <- c("#440154FF", "#7AD151FF", "#414487FF", "#fde725FF", "#2A788EFF")

# "Fishes", "Macroalgae", "Octocorals", "Scleractinian Corals", "Sponges"
# dark purple #440154
# light green #5ec962 #7AD151FF
# dark blue #3b528b #414487FF
# yellow #fde725
# teal #21918c #2A788EFF
# #25858EFF
```


## Import the database directly from the *mesophotic.org* website into R

Check its dimensions (rows x columns), and list the available fields/columns:
```{r}
# automatically download the latest database
publs <- read.csv("http://www.mesophotic.org/publications.csv", as.is=TRUE)

setwd("~/Documents/Collaborations/Mesophotic.org")

# if importing from a directory on your computer
#publs <- read.csv("publications.csv", as.is=TRUE)

publs[sapply(publs, is.character)] <- lapply(publs[sapply(publs, is.character)], as.factor)

# when need to use year as factor
publs$year <- factor(publs$publication_year)

# keep this as integer (numeric)
#publs$publication_year <- factor(publs$publication_year)

dim(publs)
```

```{r}
colnames(publs)
```

```{r}
publs %>% 
  group_by(publication_format) %>% 
   dplyr::summarise(count=n())
```

```{r}
publs %>% 
  group_by(publication_type) %>% 
   dplyr::summarise(count=n())
```

```{r}
publs %>% 
  group_by(publication_type, publication_format) %>% 
   dplyr::summarise(count=n())
```

```{r}
publs %>% 
  filter(publication_type == "scientific" & publication_format == "article" | publication_format == "chapter") %>% 
    group_by(publication_type, publication_format, validated) %>% 
      dplyr::summarise(count=n())
```

```{r}
publs %>% 
  filter(publication_type == "scientific" & publication_format == "article") %>% 
    group_by(year) %>% 
      dplyr::summarise(count=n()) %>%
          arrange(desc(count)) #%>% 
            #head()
```

#### Long-term monitoring (articles and reports) - by focusgroups
```{r}
publs_LTM_fields <- publs %>%
  filter(grepl("Long-term monitoring", fields))

publs_LTM_fields %>%
  #filter(fields == "Long-term monitoring") %>%
    group_by(publication_format) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))

# articles and reports only
publs_LTM_fields <- publs_LTM_fields %>%
  filter(publication_format != "review" & publication_format != "chapter")
```

```{r}
# Split `locations` into new rows
publs_LTM_fields_locations <- separate_rows(publs_LTM_fields, "locations", sep=";\\s+") 

# Remove rows with no associated `locations`
publs_LTM_fields_locations <- publs_LTM_fields_locations[!
(is.na(publs_LTM_fields_locations$locations) |
publs_LTM_fields_locations$locations ==
""), ]

publs_LTM_fields_locations$locations <- factor(publs_LTM_fields_locations$locations)

publs_LTM_fields_locations %>% 
    group_by(locations, publication_format) %>% 
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count))
```

```{r}
# Split `fields` into new rows
publs_LTM_fields <- separate_rows(publs_LTM_fields, "fields", sep=";\\s+") 

# Remove rows with no associated `fields`
publs_LTM_fields <- publs_LTM_fields[!
(is.na(publs_LTM_fields$fields) |
publs_LTM_fields$fields ==
""), ]

publs_LTM_fields$fields <- factor(publs_LTM_fields$fields)
```


Look at what other fields are studied in publications with Long-term monitoring
```{r}
publs_LTM_fields %>%
    group_by(fields) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))
```

Look at what focusgroups are studied in publications with Long-term monitoring
```{r}
publs_LTM_focusgroups <- publs %>%
  filter(grepl("Long-term monitoring", fields))

publs_LTM_focusgroups <- publs_LTM_focusgroups %>%
  filter(publication_format != "review" & publication_format != "chapter")

# Split `focusgroups` into new rows
publs_LTM_focusgroups <- separate_rows(publs_LTM_focusgroups, "focusgroups", sep=";\\s+") 

# Remove rows with no associated `focusgroups`
publs_LTM_focusgroups <- publs_LTM_focusgroups[!
(is.na(publs_LTM_focusgroups$focusgroups) |
publs_LTM_focusgroups$focusgroups ==
""), ]

publs_LTM_focusgroups$focusgroups <- factor(publs_LTM_focusgroups$focusgroups)

publs_LTM_focusgroups %>%
    group_by(focusgroups) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))
```


### Technical reports
```{r}
reports <- subset(publs, publication_type == "technical")
```

```{r}
# Split `locations` into new rows
reports_locations <- separate_rows(reports, "locations", sep=";\\s+") 

# Remove rows with no associated `locations`
reports_locations <- reports_locations[!
(is.na(reports_locations$locations) |
reports_locations$locations ==
""), ]

reports_locations$locations <- factor(reports_locations$locations)
```

##### Locations of reports
```{r}
reports_locations %>% 
    group_by(locations) %>% 
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count))
```


### Select only "Scientific" "Article" (original research)
```{r}
publs <- subset(publs, publication_type == "scientific" & publication_format == "article")
publs <- droplevels(publs)
```


### Select validated publications

Subset the publication list to those that are validated (i.e. metadata validated by two
different content editors) and those that are included_in_stats (i.e. as validated but only
including publications representing (peer-reviewed) scientific articles presenting original data
from mesophotic depths):
```{r}
publs <- subset(publs, validated == "true") #publs_validated
publs <- subset(publs, included_in_stats == "true") #publs_for_stats # Original data
```


##### Consider MCE and TME separately

MCE
```{r}
publs.mce <- subset(publs, mce == "true")
publs.mce <- droplevels(publs.mce)
dim(publs.mce)
```

TME
```{r}
publs.tme <- subset(publs, tme == "true")
publs.tme <- droplevels(publs.tme)
dim(publs.tme)
```


# Data analysis

## Change in publication numbers over time

Calculate the fold-increase in publications 
2008-2023
```{r}
publs_time_counts <- publs %>% #publs_validated
  group_by(publication_year) %>%
    dplyr::summarise(count=n()) %>%
      mutate(total_count = cumsum(count))

total_count <- publs_time_counts[publs_time_counts$publication_year==2023, "total_count"]/publs_time_counts[publs_time_counts$publication_year==2008, "total_count"]

total_count
```

MCE specific
```{r}
publs_time_counts <- publs.mce %>% #publs_validated
  group_by(publication_year) %>%
    dplyr::summarise(count=n()) %>%
      mutate(total_count = cumsum(count))

total_count <- publs_time_counts[publs_time_counts$publication_year==2023, "total_count"]/publs_time_counts[publs_time_counts$publication_year==2008, "total_count"]

total_count
```

TME specific
```{r}
publs_time_counts <- publs.tme %>% #publs_validated
  group_by(publication_year) %>%
    dplyr::summarise(count=n()) %>%
      mutate(total_count = cumsum(count))

total_count <- publs_time_counts[publs_time_counts$publication_year==2023, "total_count"]/publs_time_counts[publs_time_counts$publication_year==2008, "total_count"]

total_count
```

### Publications by focus groups

Where fields have multiple values (through so-called “HABTM” associations), such as
research_fields , they are separated by semicolons, for example: Ecology; Community
structure; Disturbances. 

Therefore, to query any of these fields, we need to separate these into individual rows. Then, we can plot for example the number of publications for each of those fields.

```{r}
# Split `focusgroups` into new rows
publs_focusgroups <- separate_rows(publs, "focusgroups", sep=";\\s+") #publs_for_stats

# Remove rows with no associated `focusgroups`
publs_focusgroups <- publs_focusgroups[!
(is.na(publs_focusgroups$focusgroups) |
publs_focusgroups$focusgroups ==
""), ]

publs_focusgroups$focusgroups <- factor(publs_focusgroups$focusgroups)
```

141 publications (scientific articles) with new species
*Scleractinia (Hard Corals) 3 articles, for a total of 4 new species*
```{r}
publs_focusgroups %>% 
  filter(new_species == "true") %>% 
    group_by(focusgroups, publication_format) %>%  
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count))
```

```{r}
publs_focusgroups %>%
    group_by(focusgroups) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))
```

```{r}
publs_focusgroups_counts <- publs_focusgroups %>%
     group_by(focusgroups, publication_year) %>%
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count))
```

### Corals vs. Fishes vs. Octocorals

```{r}
publs_focusgroups_counts %>% 
  filter(focusgroups == "Scleractinia (Hard Corals)" | focusgroups == "Fishes" | focusgroups == "Octocorallia (Soft Corals)") %>% # 
    ggplot(aes(x=publication_year, y=count, color = focusgroups)) +
      geom_line() + 
        geom_point() +
          #geom_area(colour="black", size=.3, alpha=.4, position = "identity") +
           scale_x_continuous(name = "Year", breaks = seq(1970, 2022, 4), limits = c(1970, 2022)) + #breaks = seq(1970, 2022, 4), limits = c(1970, 2022)
            #scale_y_continuous(limits = c(0,15), breaks = c(0,2,4,6,8,10,12,14)) +
            theme(panel.background = element_rect(fill = "white", colour = "black"), legend.position= "bottom")
```

## Publications by 'Fields'

```{r}
# Split `fields` into new rows
publs_fields <- separate_rows(publs, "fields", sep=";\\s+") #publs_for_stats

# Remove rows with no associated `fields`
publs_fields <- publs_fields[!
(is.na(publs_fields$fields) |
publs_fields$fields ==
""), ]

publs_fields$fields <- factor(publs_fields$fields)
```

```{r}
publs_fields %>%
    group_by(fields) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))
```


```{r}
publs_fields %>%
  filter(fields == "Taxonomy") %>% 
    group_by(publication_year) %>% 
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count)) %>% 
          head()
```

```{r}
publs_fields %>%
  filter(fields == "Physiology") %>% 
    group_by(publication_year) %>% 
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count)) %>% 
          head()
```


```{r}
publs_fields %>%
  filter(fields == "Molecular ecology") %>% 
    group_by(publication_year) %>% 
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count)) %>% 
          head()
```

```{r}
publs_fields %>%
  filter(fields == "Management and Conservation") %>% 
    group_by(publication_year) %>% 
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count)) %>% 
          head()
```

```{r}
publs_fields %>%
  filter(fields == "Geomorphology") %>% 
    group_by(publication_year) %>% 
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count)) %>% 
          head()
```

```{r}
publs_fields %>%
  filter(fields == "Disturbances") %>% 
    group_by(publication_year) %>% 
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count)) %>% 
          head()
```
		
```{r}
publs_fields %>%
  filter(fields == "Reproduction") %>% 
    group_by(publication_year) %>% 
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count)) %>% 
          head()
```


Within Fields, split focusgroups
```{r}
# Split `fields` into new rows
publs_fields_focusgroups <- separate_rows(publs_fields, "focusgroups", sep=";\\s+") #publs_for_stats

# Remove rows with no associated `fields`
publs_fields_focusgroups <- publs_fields_focusgroups[!
(is.na(publs_fields_focusgroups$fields) |
publs_fields_focusgroups$fields ==
""), ]

publs_fields_focusgroups$focusgroups <- factor(publs_fields_focusgroups$focusgroups)
```

```{r}
publs_fields_focusgroups %>% 
  filter(fields == "Taxonomy") %>% 
    group_by(focusgroups) %>%  
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count))
```

```{r}
publs_fields_focusgroups %>% 
  filter(fields == "Taxonomy") %>% 
    group_by(focusgroups, new_species) %>%  
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count))
```


```{r}
publs_fields_focusgroups %>%
  filter(fields == "Physiology") %>% 
    group_by(focusgroups) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))
```

```{r}
publs_fields_focusgroups %>%
  filter(fields == "Connectivity") %>% 
    group_by(focusgroups) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))
```


```{r}
publs_fields_focusgroups %>%
  filter(focusgroups == "Scleractinia (Hard Corals)" | focusgroups == "Fishes" | focusgroups ==  "Symbiodinium (zooxanthellae)") %>% 
    group_by(focusgroups, fields) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))
```

All fields
```{r}
publs_fields %>%
    group_by(fields) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))
```


Emerging fields
```{r}
publs_fields %>%
  filter(fields == "Connectivity" | fields == "Taxonomy" | fields == "Management and Conservation" | fields == "Physiology" | fields == "Geomorphology" | fields == "Disturbances" | fields == "Reproduction" | fields == "Molecular ecology" | fields == "Long-term monitoring") %>% 
    group_by(fields) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))
```


```{r}
# publs_fields$fields <- factor(publs_fields$fields, levels = c("Biodiversity", "Physiology", "Molecular ecology", "Management and Conservation", "Geomorphology", "Disturbances", "Connectivity", "Reproduction"))

publs_fields$fields <- factor(publs_fields$fields, levels = c("Reproduction", "Long-term monitoring", "Connectivity", "Disturbances", "Geomorphology", "Management and Conservation", "Physiology", "Molecular ecology", "Taxonomy"))

ridgeplot_fields <- publs_fields %>% 
  filter(fields == "Taxonomy" | fields == "Molecular ecology" | fields == "Physiology" | fields == "Management and Conservation" | fields == "Geomorphology" | fields == "Disturbances" | fields == "Connectivity" | fields == "Long-term monitoring" | fields == "Reproduction") %>%  #fields ==  "Biodiversity" | 
  ggplot(aes(x=publication_year, y=fields, fill = fields)) +
    geom_density_ridges(alpha = 0.2) +
    scale_x_continuous(name = "Year", breaks = seq(1989, 2023, 3), limits = c(1989, 2023)) +
    xlab("Year") +
    ylab("") +
    theme_minimal() +  
    theme(panel.background = element_rect(colour = "grey50"), legend.position = "none")

ridgeplot_fields
```

alternate version b - with vertical line 2008 (to show last 15 years)
```{r}
# publs_fields$fields <- factor(publs_fields$fields, levels = c("Biodiversity", "Physiology", "Molecular ecology", "Management and Conservation", "Geomorphology", "Disturbances", "Connectivity", "Reproduction"))

publs_fields$fields <- factor(publs_fields$fields, levels = c("Reproduction", "Long-term monitoring", "Connectivity", "Disturbances", "Geomorphology", "Management and Conservation", "Physiology", "Molecular ecology", "Taxonomy"))

ridgeplot_fields_b <- publs_fields %>% 
  filter(fields == "Taxonomy" | fields == "Molecular ecology" | fields == "Physiology" | fields == "Management and Conservation" | fields == "Geomorphology" | fields == "Disturbances" | fields == "Connectivity" | fields == "Long-term monitoring" | fields == "Reproduction") %>%  #fields ==  "Biodiversity" | 
  ggplot(aes(x=publication_year, y=fields, fill = fields)) +
    geom_density_ridges(alpha = 0.2) +
    geom_vline(xintercept = 2008, linetype = 2, colour = "azure4", alpha = 1) +
    scale_x_continuous(name = "Year", breaks = seq(1989, 2023, 3), limits = c(1989, 2023)) +
    xlab("Year") +
    ylab("") +
    theme_minimal() +  
    theme(panel.background = element_rect(colour = "grey50"), legend.position = "none")

ridgeplot_fields_b
```


alternate version c - with histograms
```{r}
# publs_fields$fields <- factor(publs_fields$fields, levels = c("Biodiversity", "Physiology", "Molecular ecology", "Management and Conservation", "Geomorphology", "Disturbances", "Connectivity", "Reproduction"))

publs_fields$fields <- factor(publs_fields$fields, levels = c("Reproduction", "Long-term monitoring", "Connectivity", "Disturbances", "Geomorphology", "Management and Conservation", "Physiology", "Molecular ecology", "Taxonomy"))

ridgeplot_fields_c <- publs_fields %>% 
  filter(fields == "Taxonomy" | fields == "Molecular ecology" | fields == "Physiology" | fields == "Management and Conservation" | fields == "Geomorphology" | fields == "Disturbances" | fields == "Connectivity" | fields == "Long-term monitoring" | fields == "Reproduction") %>%  #fields ==  "Biodiversity" | 
  ggplot(aes(x=publication_year, y=fields, fill = fields)) +
    geom_density_ridges(alpha = 0.2, stat="binline", bins=40) +
    scale_x_continuous(name = "Year", breaks = seq(1989, 2023, 3), limits = c(1989, 2023)) +
    xlab("Year") +
    ylab("") +
    theme_minimal() +  
    theme(panel.background = element_rect(colour = "grey50"), legend.position = "none") #panel.spacing = unit(0.1, "lines"),

ridgeplot_fields_c
```


```{r}
publs_fields_CoralFishOctocoral <- publs_fields %>% 
  filter(focusgroups == "Scleractinia (Hard Corals)" | focusgroups == "Fishes" | focusgroups == "Octocorallia (Soft Corals)")
```


```{r}
#publs_fields_CoralFishOctocoral$focusgroups <- recode(publs_fields_CoralFishOctocoral$focusgroups, 'Algae (Macro, Turf and Crustose Coralline)' = "Macroalgae")

publs_fields_CoralFishOctocoral$focusgroups <- recode(publs_fields_CoralFishOctocoral$focusgroups, 'Octocorallia (Soft Corals)' = "Octocorals")

#publs_fields_CoralFishOctocoral$focusgroups <- recode(publs_fields_CoralFishOctocoral$focusgroups, 'Porifera (Sponges)' = "Sponges")

publs_fields_CoralFishOctocoral$focusgroups <- recode(publs_fields_CoralFishOctocoral$focusgroups, 'Scleractinia (Hard Corals)' = "Scleractinian corals")

publs_fields_CoralFishOctocoral$focusgroups <- factor(publs_fields_CoralFishOctocoral$focusgroups, levels = c("Octocorals", "Fishes", "Scleractinian corals"))

barplot_fields <- publs_fields_CoralFishOctocoral %>% 
  filter(fields == "Taxonomy" | fields == "Molecular ecology" | fields == "Physiology" | fields == "Management and Conservation" | fields == "Geomorphology" | fields == "Disturbances" | fields == "Connectivity" | fields == "Long-term monitoring" | fields == "Reproduction") %>%
  group_by(fields, focusgroups) %>% 
      dplyr::summarise(count=n()) %>%
        ggplot(aes(x = count, y = fields, fill = focusgroups)) +
          geom_bar(stat="identity", position="dodge") + #, fill = "grey"
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          #scale_fill_viridis(discrete=TRUE, name="") +
          scale_fill_manual(values=c("#25858EFF", "#440154", "#fde725")) +
          scale_x_continuous(limits = c(0, 60)) +
          xlab("Publications") +
          #ylab("") +
          #labs(x = "Percent", y = "Fields", title = "TME") +
           theme_minimal() +  
           theme(panel.background = element_rect(colour = "grey50"), axis.text.y=element_blank(), axis.title.y=element_blank(), legend.key.height = unit(1, "null"), legend.position = c(.7,.05), legend.title = element_text(size = 6), legend.text = element_text(size = 6))

# legend.key.width = unit(5, "null"), 
# By setting the legend.key{.height/.width} theme argument as a "null" unit, legends can now expand to fill the available space.

barplot_fields
```

Break out into top focusgroups - Scleractinian corals, Fishes, Octocorallia



merge ridgeline with barplot counts
```{r}
# cowplot

#pdf("Fig Fields emerging rideplot barplot combo_focusgroups_07-28.pdf", width = 8, height = 6, encoding="MacRoman") #width = 10

plot <- plot_grid(ridgeplot_fields, 
                  barplot_fields,
                  #labels = c('a)', 'b)', 'c)', 'd)'), 
                  #hjust = -1, 
                  #vjust = 1, 
                  #align = "hv",
                  rel_widths = c(3, 1),
                  rel_heights = 1,
                  #scale = c(1, 1. 1, 1),
                  label_size = 12,
                  ncol = 2,
                  label_fontface = "plain")

plot

#dev.off()

#svg("Fig Fields emerging rideplot barplot combo_focusgroups_07-28.svg", width = 8, height = 6)
#print(plot)
#dev.off()
```


Alternate ridgeline b with barplot counts
```{r}
# cowplot

#pdf("Fig Fields emerging rideplot barplot combo_focusgroups_07-28_ALT-2008-vline.pdf", width = 8, height = 6, encoding="MacRoman") #width = 10

plot <- plot_grid(ridgeplot_fields_b, 
                  barplot_fields,
                  #labels = c('a)', 'b)', 'c)', 'd)'), 
                  #hjust = -1, 
                  #vjust = 1, 
                  #align = "hv",
                  rel_widths = c(3, 1),
                  rel_heights = 1,
                  #scale = c(1, 1. 1, 1),
                  label_size = 12,
                  ncol = 2,
                  label_fontface = "plain")

plot

#dev.off()

#svg("Fig Fields emerging rideplot barplot combo_focusgroups_07-28_ALT-2008-vline.svg", width = 8, height = 6)
#print(plot)
#dev.off()
```

## Publications by 'Platforms'
```{r}
# Split `platforms` into new rows
publs_platforms <- separate_rows(publs, "platforms", sep=";\\s+") #publs_for_stats

# Remove rows with no associated `platforms`
publs_platforms <- publs_platforms[!
(is.na(publs_platforms$platforms) |
publs_platforms$platforms ==
""), ]

publs_platforms$platforms <- factor(publs_platforms$platforms)

#write.csv(publs_platforms, "publs_platforms.csv")
```


```{r}
publs_platforms %>%
     group_by(platforms) %>%
      dplyr::summarise(count=n()) %>% 
        arrange(desc(count))
```


### Platforms by Location

```{r}
# Split `locations` into new rows
publs_platforms_location <- separate_rows(publs_platforms, "locations", sep=";\\s+") #publs_for_stats

# Remove rows with no associated `locations`
publs_platforms_location <- publs_platforms_location[!
(is.na(publs_platforms_location$locations) |
publs_platforms_location$locations ==
""), ]

publs_platforms_location$locations <- factor(publs_platforms_location$locations)

#write.csv(publs_platforms_location, "publs_platforms_location.csv")
```


```{r}
publs_platforms_location %>%
  filter(platforms == "Manned Submersible") %>% 
    group_by(locations) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))
```

```{r}
publs_platforms_location %>%
  filter(platforms == "Rebreather") %>% 
    group_by(locations) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))
```


Platforms used by location - organized by platform
```{r}
publs_platforms_location_counts <- publs_platforms_location %>%
    group_by(platforms, locations) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(15, count) %>%
          arrange(desc(count))

head(publs_platforms_location_counts)
```

top 10
```{r}
publs_platforms_location %>%
    group_by(platforms, locations) %>% 
      dplyr::summarise(count=n()) %>% 
        top_n(10, count) %>%
          arrange(desc(count))
```


## Publications by 'Locations'

##### Consider MCE and TME separately

MCE - publications that only had MCE locations
```{r}
publs.mce <- subset(publs, mce == "true" & tme == "false")
publs.mce <- droplevels(publs.mce)
```

```{r}
# Split `locations` into new rows
publs.mce_locations <- separate_rows(publs.mce, "locations", sep=";\\s+") #publs_for_stats

# Remove rows with no associated `locations`
publs.mce_locations <- publs.mce_locations[!
(is.na(publs.mce_locations$locations) |
publs.mce_locations$locations ==
""), ]

publs.mce_locations$locations <- factor(publs.mce_locations$locations)
```

```{r}
n_distinct(publs.mce_locations$locations)
#unique(publs.mce_locations$locations)
```

```{r}
#Separate by periods, before 2008 and 2008-2023
publs.mce_locations$range <- ifelse(publs.mce_locations$publication_year <= 2007, '<2008',
                       ifelse(publs.mce_locations$publication_year <=2023, '<2023', 'unassigned'))
```

```{r}
publs.mce_locations %>% 
  group_by(range) %>% 
    dplyr::summarise(count=n())
```

```{r}
#before 2008
publs_mce_2008 <- subset(publs.mce_locations, range == "<2008")
#publs_mce_2008 <- droplevels(publs_mce_2008)
n_distinct(publs_mce_2008$locations)
```

```{r}
publs_mce_2008 %>% 
  group_by(locations) %>% 
    dplyr::summarise() %>% #dplyr::summarise(count=n())
      head()
```

```{r}
#2008-2023
publs_mce_2023 <- subset(publs.mce_locations, range == "<2023")
n_distinct(publs_mce_2023$locations)
```

```{r}
publs_mce_2023 %>% 
  group_by(locations) %>% 
    dplyr::summarise() %>% #dplyr::summarise(count=n())
      head()
```

*However, there is overlap of locations that are present in both Before 2008 and 2008-2023 data*

***Cross check to only obtain unique locations***

In summary, 51 MCE locations were both in Before 2008 and 2008-2023 categories
Therefore 98 new locations 2008-2023 - 51 shared locations = 47 new MCE locations between 2008-2023
Note: 14 unique MCE locations that were only from Before 2008 (but not studied again in 2008-2023)
Grand total = 14 unique locations Before 2008 + 51 shared locations two time periods + 47 new MCE locations 2008-2023 = TOTAL 112 unique MCE locations
```{r}
#find duplicates between two data frames
dplyr::intersect(publs_mce_2008$locations, publs_mce_2023$locations)
```

Number of publications per location
```{r}
publs.mce_locations_counts <- publs.mce_locations %>% 
  group_by(locations) %>% 
    dplyr::summarise(count=n()) %>% 
      arrange(count)

publs.mce_locations_counts

#write.csv(publs.mce_locations_counts, "publs.mce_locations_counts.csv")
```


#### TME - publications that only had TME locations
```{r}
publs.tme <- subset(publs, tme == "true" & mce == "false")
publs.tme <- droplevels(publs.tme)
```

```{r}
# Split `locations` into new rows
publs.tme_locations <- separate_rows(publs.tme, "locations", sep=";\\s+") #publs_for_stats

# Remove rows with no associated `locations`
publs.tme_locations <- publs.tme_locations[!
(is.na(publs.tme_locations$locations) |
publs.tme_locations$locations ==
""), ]

publs.tme_locations$locations <- factor(publs.tme_locations$locations)
```

```{r}
n_distinct(publs.tme_locations$locations)
#unique(publs.tme_locations$locations)
```

```{r}
#Separate by periods, before 2008 and 2008-2023
publs.tme_locations$range <- ifelse(publs.tme_locations$publication_year <= 2007, '<2008',
                       ifelse(publs.tme_locations$publication_year <=2023, '<2023', 'unassigned'))
```

```{r}
publs.tme_locations %>% 
  group_by(range) %>% 
    dplyr::summarise(count=n())
```

```{r}
#before 2008
publs_tme_2008 <- subset(publs.tme_locations, range == "<2008")
n_distinct(publs_tme_2008$locations)
```

```{r}
#2008-2023
publs_tme_2023 <- subset(publs.tme_locations, range == "<2023")
n_distinct(publs_tme_2023$locations)
```

In summary, 13 TME locations were both in Before 2008 and 2008-2023 categories
Therefore 44 new locations 2008-2023 - 13 shared locations = 31 new TME locations between 2008-2023
Therefore, 14 Before 2008 locations + 31 new 2008-2023 locations = 45 locations total
```{r}
#find duplicates between two data frames
dplyr::intersect(publs_tme_2008$locations, publs_tme_2023$locations)
```


Number of publications per location
```{r}
publs.tme_locations_counts <- publs.tme_locations %>% 
  group_by(locations) %>% 
    dplyr::summarise(count=n()) %>% 
      arrange(count)

publs.tme_locations_counts

#write.csv(publs.tme_locations_counts, "publs.tme_locations_counts.csv")
```


### Overall locations
```{r}
# Split `locations` into new rows
publs_locations <- separate_rows(publs, "locations", sep=";\\s+") #publs_for_stats

# Remove rows with no associated `locations`
publs_locations <- publs_locations[!
(is.na(publs_locations$locations) |
publs_locations$locations ==
""), ]

publs_locations$locations <- factor(publs_locations$locations)
```

```{r}
#Separate by periods, before 2008 and 2008-2023
publs_locations$range <- ifelse(publs_locations$publication_year <= 2007, '<2008',
                       ifelse(publs_locations$publication_year <=2023, '<2023', 'unassigned'))
```

```{r}
#before 2008
publs_2008 <- subset(publs_locations, range == "<2008")
n_distinct(publs_2008$locations)
```

```{r}
#2008-2023
publs_2023 <- subset(publs_locations, range == "<2023")
n_distinct(publs_2023$locations)
```

```{r}
publs_locations %>%
    group_by(locations) %>% 
      dplyr::summarise(count=n()) %>% 
        #top_n(10, count) %>%
          arrange(desc(count))
```


```{r}
publs_locations <- subset(publs_locations, original_data == "true")

#write.csv(publs_locations, "publs_locations_scientific_articles_original-data.csv")
```




